The role of cognateness in non-native spoken word recognition

### 2022-01-14 docs/report.Rmd
update figures

* Aged 18 to 25 years
* Did not report being fluent in Catalan, Spanish, or Italian
* Provided at least 80% of the presented trials
* Did not report motor, auditory or visual (other than wearing glasses) problems

@@

```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  count(test_language, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
  geom_col(position = position_dodge()) +
  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

>>>>

* Aged 18 to 26 years
* Did not report being fluent in Catalan, Spanish, or Italian
* Provided at least 80% of the presented trials
* Did not report motor, auditory or visual (other than wearing glasses) problems

@@

```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  count(group, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = group)) +
  facet_wrap(~group, ncol = 1) +
  geom_col() +
  labs(x = "number of phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

### 2022-01-14 manuscript/manuscript.Rmd
update figures

abstract: |
  When listening to non-native speech, words whose translation in the native language is phonologically similar are easier to recognise. Previous studies suggest that low-proficiency bilinguals benefit more from this facilitation effect than high-proficiency bilinguals, but this claim has remained relatively unaddressed. In this study, we tested English and Spanish monolinguals in a translation elicitation task, in which they had to guess the translation of a list of words in an unfamiliar language (Spanish or Catalan). Presented words varied in their phonological similarity to their correct translation. This allowed us to test the impact of such phonological similarity on word recognition in a sample of listeners with very low proficiency in a non-native language. Translation accuracy increased substantially the likelihood of correct translations, but this effect was mediated by the size of the phonological neighbourhood of the correct translation. Participants benefitted from phonological similarity only when the correct translation was part of a sparse phonological neighbourhood. When the correct translation was part of a larger phonological neighbourhood, translation accuracy remained unaffected. Overall, these results provide evidence that litening to non-native words activated lexical representations in the native language, even in monolingual participants, resulting in competition for selection in the native language during translation.

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla). After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Then, participants completed the experimental task. Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials. Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio ended. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word. Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish). The task took approximately 15 minutes to be completed.

@@

We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist]. This function computes the inverse of the Levenshtein distance between two character strings as a proportion. First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary]. This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1. The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical. We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists. Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.

```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity. Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours. The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds). The diagonal of the plot matrix indicates the marginal distribution of eacch variable. The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
my_bin <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_density(..., size = 1, alpha = 0.25) +
    geom_rug(...)
}

my_smooth <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
}

my_cor <- function(data, mapping, ...) {
  ggplot(data = data, mapping = mapping) +
    ggally_cor(..., size = 7, title = "r", )
}

stimuli %>% 
  drop_na(frequency_zipf, pthn, lv) %>% 
  ggpairs(
    mapping = aes(colour = group, fill = group),
    columns = c("frequency_zipf", "pthn", "lv"),
    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
    lower = list(continuous = my_smooth),
    diag = list(continuous = my_bin),
    upper = list("my_cor")
  ) +
  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)

@@

```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

tb_stimuli <- stimuli %>% 
  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
  group_by(group) %>% 
  summarise_at(
    vars(frequency:lv), 
    list(
      mean = ~mean(., na.rm = TRUE),  
      sd = ~sd(., na.rm = TRUE)
    )
  ) %>%
  gt(rowname_col = "group") %>% 
  tab_header(
    title = md("Table 1"),
    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
  ) %>% 
  tab_source_note(
    md("*Note*. Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
  ) %>% 
  fmt_number(frequency_mean:pthn_sd) %>%
  fmt_percent(matches("lv")) %>% 
  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
  cols_label(
    frequency_mean = md("Mean"),
    frequency_sd = md("*SD*"),
    frequency_zipf_mean = md("Mean"),
    frequency_zipf_sd = md("*SD*"),
    pthn_mean = md("Mean"),
    pthn_sd = md("*SD*"),
    lv_mean = md("Mean"),
    lv_sd = md("*SD*")
  ) %>% 
  summary_rows(
    columns = c(matches("mean"), -matches("lv")),
    fns = list(
      Mean = "mean",
      SD = "sd"
    )
  ) %>% 
  summary_rows(
    columns = matches("lv"),
    fns = list(
      Mean = "mean",
      SD = "sd"
    ), formatter = fmt_percent
  ) %>% 
  tab_style(
    style = cell_text(align = "left", weight = "bold", size = "medium"),
    locations = cells_title(groups = "title")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium", style = "italic"),
    locations = cells_title(groups = "subtitle")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium"),
    locations = cells_source_notes()
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium"),
    locations = cells_stub_grand_summary()
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(), cells_column_spanners(), 
      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = 1:9)
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:9)
  ) %>% 
  gtsave(here("Figures", "stimuli.png"))

include_graphics(here("Figures", "stimuli.png"), auto_pdf = FALSE)

@@

We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution. We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0). Second, we added the fixed effect of lexical frequency of the target word ($Frequency$). We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction). We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.5) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize]. Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random]. We will later refer to these models as Model 0 to Model 6. For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.

@@

tab_loos <- model_loos %>% 
  as.data.frame() %>% 
  rownames_to_column("model") %>% 
  arrange(desc(elpd_loo)) %>% 
  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
  mutate_all(~ifelse(. == 0, NA, .)) %>% 
  gt(rowname_col = "model") %>% 
  tab_header(
    title = md("Table 2"),
    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
  ) %>% 
  tab_source_note(
    md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
  ) %>% 
  fmt_missing(everything(), missing_text = "-") %>% 
  fmt_number(2:7) %>% 
  cols_label(
    model = md("**Model**"),
    looic = md("***LOO<sub>IC</sub>***"),
    se_looic = md("***SE*<sub>IC</sub>**"),
    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
    se_elpd_loo = md("***SE***"),
    elpd_diff = md("***LOO<sub>diff</sub>***"),
    se_diff = md("***SE<sub>diff</sub>***")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", weight = "bold", size = "medium"),
    locations = cells_title(groups = "title")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium", style = "italic"),
    locations = cells_title(groups = "subtitle")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium"),
    locations = cells_source_notes()
  ) %>% 
  tab_style(
    style = list(cell_borders(sides = c("left", "right"), color = "white"),
                 cell_text(align = "center")),
    locations = list(
      cells_body(), cells_row_groups(), cells_column_spanners(), 
      cells_stub(), cells_source_notes(),
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "medium"), 
                 cell_borders(sides = "all", color = "white")),
    locations = list(cells_stub(), cells_source_notes())
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = 1:7)
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:7)
  ) %>% 
  gtsave(here("Figures", "loos.png"))

include_graphics(here("Figures", "loos.png"))

@@

  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
  stat_pointinterval(.width = 0, point_size = 1) +
  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +

@@

The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data]. Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_Intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_Intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]). Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants. Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`. The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]). This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.

>>>>

abstract: |

@@

The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2]. Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla). After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Then, participants completed the experimental task. Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials. Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio ended. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word. Trials in which participants took longer than 10 seconds to respond were also filtered out. Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish). The task took approximately 15 minutes to be completed.

@@

```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.cap="Distribution and summary statistics of stimuli properties. A\\) Lexical frequency of target words, expressed in counts per million and as Zipf scores. B\\) Number of higher-frequency phonological neighbourhoods of target words (PTHN). C\\) Normalised Levenshtein phonological similarity between the presented and the target words, expressed in percentages."}

medians <- stimuli %>% 
  group_by(group) %>% 
  summarise(
    pthn_median = median(pthn),
    pthn_sd = sd(pthn),
    frequency_zipf_median = median(frequency_zipf),
    frequency_zipf_sd = sd(frequency_zipf),
    lv_median = median(lv),
    lv_sd = sd(lv)
  )

stimuli %>% 
  count(group, pthn) %>% 
  ggplot(aes(pthn, n, fill = group)) + 
  facet_wrap(~group, ncol = 1) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = 0, xmax = pthn_median+pthn_sd, 
      ymin = -Inf, ymax = Inf, fill = group
    ), inherit.aes = FALSE, alpha = 0.15
  ) +   
  geom_col(colour = "white") + 
  geom_vline(
    data = medians, aes(xintercept = pthn_median, group = group), 
    colour = "black", size = 0.5, linetype = "dashed"
  ) + 
  geom_text(data = medians, aes(label = "Median ± SD", x = pthn_median+10, y = 40), size = 3) +
  geom_curve(
    data = medians, aes(x = pthn_median+10, xend = pthn_median+0.2, y = 35, yend = 35),
    arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30") +
  labs(x = "PTHN", y = "Number of trials", title = "Higher-frequency neighbours") +

  ggplot(stimuli, aes(frequency_zipf, fill = group)) + 
  facet_wrap(~group, ncol = 1) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
      ymin = -Inf, ymax = Inf, fill = group
    ), inherit.aes = FALSE, alpha = 0.15
  ) +   
  geom_histogram(bins = 20, colour = "white") +
  geom_vline(
    data = medians, aes(xintercept = frequency_zipf_median, group = group), 
    colour = "black", size = 0.5, linetype = "dashed"
  ) + 
  labs(x = "Lexical frequency (Zipf)", y = "Number of trials", title = "Frequency") +

  ggplot(stimuli, aes(lv, fill = group)) +
  facet_wrap(~group, ncol = 1) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = 0, xmax = lv_median+lv_sd, 
      ymin = -Inf, ymax = Inf, fill = group
    ), inherit.aes = FALSE, alpha = 0.15
  ) +   
  geom_histogram(bins = 20, colour = "white") +
  geom_vline(
    data = medians, aes(xintercept = lv_median, group = group), 
    colour = "black", size = 0.5, linetype = "dashed"
  ) + 
  scale_x_continuous(labels = percent) +
  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity") +

  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "A") &
  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
  theme(legend.position = "none")
```

We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist]. This function computes the inverse of the Levenshtein distance between two character strings as a proportion. First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary]. This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1. The result is percentage score indicating no similarity between the two strings if 0%, and a perfect match between two strings if 100%. We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists. Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.

@@

@@

We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution. We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0). Second, we added the fixed effect of lexical frequency of the target word ($Frequency$). We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction). We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.25) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize]. Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random]. We will later refer to these models as Model 0 to Model 6. For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.

@@

tab_loos <- model_loos %>% 
  as.data.frame() %>% 
  rownames_to_column("model") %>% 
  arrange(desc(elpd_loo)) %>% 
  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
  mutate_all(~ifelse(. == 0, NA, .)) %>% 
  gt(rowname_col = "model") %>% 
  tab_header(
    title = md("Table 2"),
    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
  ) %>% 
  tab_source_note(
    md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
  ) %>% 
  fmt_missing(everything(), missing_text = "-") %>% 
  fmt_number(2:7) %>% 
  cols_label(
    model = md("**Model**"),
    looic = md("***LOO<sub>IC</sub>***"),
    se_looic = md("***SE*<sub>IC</sub>**"),
    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
    se_elpd_loo = md("***SE***"),
    elpd_diff = md("***LOO<sub>diff</sub>***"),
    se_diff = md("***SE<sub>diff</sub>***")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", weight = "bold", size = "medium"),
    locations = cells_title(groups = "title")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium", style = "italic"),
    locations = cells_title(groups = "subtitle")
  ) %>% 
  tab_style(
    style = cell_text(align = "left", size = "medium"),
    locations = cells_source_notes()
  ) %>% 
  tab_style(
    style = list(cell_borders(sides = c("left", "right"), color = "white"),
                 cell_text(align = "center")),
    locations = list(
      cells_body(), cells_row_groups(), cells_column_spanners(), 
      cells_stub(), cells_source_notes(),
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "medium"), 
                 cell_borders(sides = "all", color = "white")),
    locations = list(cells_stub(), cells_source_notes())
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = 1:7)
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:7)
  ) %>% 
  gtsave(here("img", "loos.png"))

include_graphics(here("img", "loos.png"))

@@

  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
  stat_pointinterval(.width = 0, point_size = 1) +
  scale_x_continuous(labels = percent, limits = c(0, 1)) +
  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +

@@

The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data]. Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]). Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants. Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`. The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]). This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
